---
title: "Modules"
format: html
---

```{r}
#| include: false
library(checkdown)
right_ans <- paste("Correct", emo::ji("+1"))
wrong_ans <- "That is not correct. Rewatch the video if needed"

# Use callout box for checkdown questions
practice_question <- function(qtitle, ctitle = "Check your learning", ...) {
    cat("::: {.callout-note icon=false}\n")
    cat("\n## ", ctitle, "\n")
    cat("\n", qtitle, "\n")
    print(
        check_question(right = right_ans, wrong = wrong_ans, ...)
    )
    cat("\n:::\n")
}
```

::: {.panel-tabset}

# Week 1

## Week Learning Objectives

By the end of this module, you will be able to 

- Navigate the course website and Blackboard site
- Identify the Slack channels relevant for the course
- Describe the historical origin of Bayesian statistics
- Identify components in research papers involving Bayesian analyses
- Render a simple Quarto (.qmd) file

## Task List

1. Review the [syllabus](/syllabus/index.qmd)
2. Review the resources (slides and [note][5])
3. Install/Update R and RStudio on your computer
4. Attend the Tuesday and Thursday class meetings
5. Complete the assigned readings
    * [McElreath ch. 1][2]
    * Supplemental (i.e., optional) reading: [Gigerenzer (2004)][1]
    * [Markdown Basics][4]
6. Introduce yourself on the #introduction Slack channel (as part of HW 1)
7. Complete Homework 1 (see instruction on Brightspace)

[1]: https://pure.mpg.de/rest/items/item_2101336/component/file_2101335/content
[2]: https://xcelab.net/rmpubs/sr2/statisticalrethinking2_chapters1and2.pdf
[4]: https://quarto.org/docs/authoring/markdown-basics.html
[5]: docs/01-intro.html

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/intro.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/intro.qmd){target="_blank"}

# Week 2

## Week Learning Objectives

By the end of this module, you will be able to 

- Describe the **subjectivist** interpretation of probability, and contrast it with the **frequentist** interpretation
- Compute probability density using **simulations**
- Compute **joint, marginal, and conditional probabilities** with two variables
- Apply Bayes' rule to obtain **posterior** from prior and data
- Explain what **data-order invariance** and **exchangeability** are
- Use **grid approximation** to obtain the posterior for a **Bernoulli** model

## Task List

1. Review the resources (slides and [notes][9])
2. Attend the Tuesday and Thursday class meetings
3. Complete the assigned readings
    * [McElreath ch. 2][2], 3
    * [James ch. 1][3]
4. Complete Homework 2 (due in two weeks; see instruction on Brightspace)

[2]: https://xcelab.net/rmpubs/sr2/statisticalrethinking2_chapters1and2.pdf
[3]: https://jamesstone.sites.sheffield.ac.uk/books/bayes-rule/an-introduction-to-bayes-rule-chapter-1
[9]: docs/02-probability.html

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/probability_and_bayes_theorem.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/probability_and_bayes_theorem.qmd){target="_blank"}

# Week 3

## Week Learning Objectives

By the end of this module, you will be able to 

- Apply **Bayesian workflow** to analyze real data with a **Bernoulli** model
- Explain the idea of a **conjugate prior**
- Summarize the posterior distribution using **simulations**
- Apply Bayesian **terminology** in summarizing the posterior
- Use R to perform **prior** and **posterior predictive checks**

## Task List

1. Review the resources (slides and [notes][11])
2. Attend the Tuesday and Thursday class meetings
3. Complete the assigned readings
    * [Johnson ch. 3][6]
    * [Getting started with CmdStanR][10]
4. Complete Homework 2 (see instruction on Brightspace)

[6]: https://www.bayesrulesbook.com/chapter-3
[10]: https://mc-stan.org/cmdstanr/articles/cmdstanr.html
[11]: docs/04-beta-bernoulli-model.html

## Lecture Videos

::: {.panel-tabset}

### Recap of the Beta-Bernoulli Model

{{< video https://youtu.be/TAzX24EpAT4 >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "If we do not use a conjugate prior for the Bernoulli model, what is the distribution of the posterior $\\theta$?",
    answer = "Something not from a known family of distributions", 
    options = c("Normal distribution", 
                "Beta distribution",
                "Something not from a known family of distributions"), 
    type = "radio"
)
```

### Summarizing the Posterior

{{< video https://youtu.be/qb9jDxDd4W0 >}}

```{r}
#| echo: false
library(ggplot2)
ggplot() +
    stat_function(fun = dchisq, args = list(df = 3), 
                  geom = "area", fill = "lightblue",
                  alpha = 0.5, xlim = c(qchisq(.1, df = 3), qchisq(.9, df = 3))) + 
    stat_function(fun = dchisq, args = list(df = 3)) +
    lims(x = c(0, 10))
```

```{r}
#| echo: false
#| results: asis
practice_question(
    "In the posterior distribution above, the interval bounding the shaded area is",
    answer = "an equal-tailed credible interval", 
    options = c("an equal-tailed credible interval", 
                "a highest density interval",
                "a one-sided credible interval"), 
    type = "radio"
)
```

### Posterior Predictive Check

{{< video https://youtu.be/UuVMwpnzWU0 >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "A researcher asks a participant 10 true/false questions to assess their statistics knowledge. A posterior predictive distribution in this case would be",
    answer = "a probability distribution of future true/false responses from the participant, based on the posterior distribution of their statistics knowledge", 
    options = c("a probability distribution of the knowledge of the participant, given their answers", 
                "a probability distribution of future true/false responses from the participant, based on the posterior distribution of their statistics knowledge",
                "a probability distribution of the observed responses from the participant, based on the prior distribution of their statistics knowledge",
                "a probability distribution of the knowledge of the participant, before incorporating the observed responses from them"), 
    type = "radio"
)
```

### Stan (Part 1)

{{< video https://youtu.be/TuAZR3Y2EAo >}}

### Stan (Part 2)

{{< video https://youtu.be/T2XpwCe6aJU >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "Which of the following is the correct way to declare a data variable of responses on a 5-point scale from N participants?",
    answer = "array[N] int<lower=1, upper=5> y;", 
    options = c("array[N] int<lower=1, upper=5> y", 
                "array[N] int<lower=1, upper=5> y;",
                "vector[N] y",
                "vector[N] y;"), 
    type = "radio"
)
```

### Stan (Part 3)

{{< video https://youtu.be/GQzhUeXeS0o >}}

:::

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/one_parameter_models.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/one_parameter_models.qmd){target="_blank"}

# Week 4

## Week Learning Objectives

By the end of this module, you will be able to 

- Explain the logic of a **hierarchical model**
- Apply the **binomial distribution** to describe the sum of multiple Bernoulli trials
- Program a hierarchical binomial model in Stan
- Analyze secondary data using a hierarchical normal model (i.e., random-effect meta-analysis)

## Task List

1. Review the resources (slides and [notes][17])
2. Watch the lecture videos below (to be posted)
3. Complete the assigned readings
    * McElreath ch. 13.1, 13.2
    * [Gabry et al. (2019)][12]
    * [Gelman et al. (2010)][13]
4. Start working on Homework 3 (see instruction on Brightspace)

[12]: https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssa.12378
[13]: https://arxiv.org/abs/2011.01808
[17]: docs/05-hierarchical-models.html

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/hierarchical_models.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/hierarchical_models.qmd){target="_blank"}

:::

P.S.: If'd like to print the slides to PDF, follow <https://quarto.org/docs/presentations/revealjs/presenting.html#print-to-pdf>