---
title: "Modules"
format: html
---

```{r}
#| include: false
library(checkdown)
right_ans <- paste("Correct", emo::ji("+1"))
wrong_ans <- "That is not correct. Rewatch the video if needed"

# Use callout box for checkdown questions
practice_question <- function(qtitle, ctitle = "Check your learning", ...) {
    cat("::: {.callout-note icon=false}\n")
    cat("\n## ", ctitle, "\n")
    cat("\n", qtitle, "\n")
    print(
        check_question(right = right_ans, wrong = wrong_ans, ...)
    )
    cat("\n:::\n")
}
```

::: {.panel-tabset}

# Week 1

## Week Learning Objectives

By the end of this module, you will be able to 

- Navigate the course website and Blackboard site
- Identify the Slack channels relevant for the course
- Describe the historical origin of Bayesian statistics
- Identify components in research papers involving Bayesian analyses
- Render a simple Quarto (.qmd) file

## Task List

1. Review the [syllabus](/syllabus/index.qmd)
2. Review the resources (slides and [note][5])
3. Install/Update R and RStudio on your computer
4. Attend the Tuesday and Thursday class meetings
5. Complete the assigned readings
    * [McElreath ch. 1][2]
    * Supplemental (i.e., optional) reading: [Gigerenzer (2004)][1]
    * [Markdown Basics][4]
6. Introduce yourself on the #introduction Slack channel (as part of HW 1)
7. Complete Homework 1 (see instruction on Brightspace)

[1]: https://pure.mpg.de/rest/items/item_2101336/component/file_2101335/content
[2]: https://xcelab.net/rmpubs/sr2/statisticalrethinking2_chapters1and2.pdf
[4]: https://quarto.org/docs/authoring/markdown-basics.html
[5]: docs/01-intro.html

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/intro.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/intro.qmd){target="_blank"}

# Week 2

## Week Learning Objectives

By the end of this module, you will be able to 

- Describe the **subjectivist** interpretation of probability, and contrast it with the **frequentist** interpretation
- Compute probability density using **simulations**
- Compute **joint, marginal, and conditional probabilities** with two variables
- Apply Bayes' rule to obtain **posterior** from prior and data
- Explain what **data-order invariance** and **exchangeability** are
- Use **grid approximation** to obtain the posterior for a **Bernoulli** model

## Task List

1. Review the resources (slides and [notes][9])
2. Attend the Tuesday and Thursday class meetings
3. Complete the assigned readings
    * [McElreath ch. 2][2], 3
    * [James ch. 1][3]
4. Complete Homework 2 (due in two weeks; see instruction on Brightspace)

[2]: https://xcelab.net/rmpubs/sr2/statisticalrethinking2_chapters1and2.pdf
[3]: https://jamesstone.sites.sheffield.ac.uk/books/bayes-rule/an-introduction-to-bayes-rule-chapter-1
[9]: docs/02-probability.html

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/probability_and_bayes_theorem.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/probability_and_bayes_theorem.qmd){target="_blank"}

# Week 3

## Week Learning Objectives

By the end of this module, you will be able to 

- Apply **Bayesian workflow** to analyze real data with a **Bernoulli** model
- Explain the idea of a **conjugate prior**
- Summarize the posterior distribution using **simulations**
- Apply Bayesian **terminology** in summarizing the posterior
- Use R to perform **prior** and **posterior predictive checks**

## Task List

1. Review the resources (slides and [notes][11])
2. Attend the Tuesday and Thursday class meetings
3. Complete the assigned readings
    * [Johnson ch. 3][6]
    * [Getting started with CmdStanR][10]
4. Complete Homework 2 (see instruction on Brightspace)

[6]: https://www.bayesrulesbook.com/chapter-3
[10]: https://mc-stan.org/cmdstanr/articles/cmdstanr.html
[11]: docs/04-beta-bernoulli-model.html

## Lecture Videos

::: {.panel-tabset}

### Recap of the Beta-Bernoulli Model

{{< video https://youtu.be/TAzX24EpAT4 >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "If we do not use a conjugate prior for the Bernoulli model, what is the distribution of the posterior $\\theta$?",
    answer = "Something not from a known family of distributions", 
    options = c("Normal distribution", 
                "Beta distribution",
                "Something not from a known family of distributions"), 
    type = "radio"
)
```

### Summarizing the Posterior

{{< video https://youtu.be/qb9jDxDd4W0 >}}

```{r}
#| echo: false
library(ggplot2)
ggplot() +
    stat_function(fun = dchisq, args = list(df = 3), 
                  geom = "area", fill = "lightblue",
                  alpha = 0.5, xlim = c(qchisq(.1, df = 3), qchisq(.9, df = 3))) + 
    stat_function(fun = dchisq, args = list(df = 3)) +
    lims(x = c(0, 10))
```

```{r}
#| echo: false
#| results: asis
practice_question(
    "In the posterior distribution above, the interval bounding the shaded area is",
    answer = "an equal-tailed credible interval", 
    options = c("an equal-tailed credible interval", 
                "a highest density interval",
                "a one-sided credible interval"), 
    type = "radio"
)
```

### Posterior Predictive Check

{{< video https://youtu.be/UuVMwpnzWU0 >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "A researcher asks a participant 10 true/false questions to assess their statistics knowledge. A posterior predictive distribution in this case would be",
    answer = "a probability distribution of future true/false responses from the participant, based on the posterior distribution of their statistics knowledge", 
    options = c("a probability distribution of the knowledge of the participant, given their answers", 
                "a probability distribution of future true/false responses from the participant, based on the posterior distribution of their statistics knowledge",
                "a probability distribution of the observed responses from the participant, based on the prior distribution of their statistics knowledge",
                "a probability distribution of the knowledge of the participant, before incorporating the observed responses from them"), 
    type = "radio"
)
```

### Stan (Part 1)

{{< video https://youtu.be/TuAZR3Y2EAo >}}

### Stan (Part 2)

{{< video https://youtu.be/T2XpwCe6aJU >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "Which of the following is the correct way to declare a data variable of responses on a 5-point scale from N participants?",
    answer = "array[N] int<lower=1, upper=5> y;", 
    options = c("array[N] int<lower=1, upper=5> y", 
                "array[N] int<lower=1, upper=5> y;",
                "vector[N] y",
                "vector[N] y;"), 
    type = "radio"
)
```

### Stan (Part 3)

{{< video https://youtu.be/GQzhUeXeS0o >}}

:::

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/one_parameter_models.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/one_parameter_models.qmd){target="_blank"}

# Week 4

## Week Learning Objectives

By the end of this module, you will be able to 

- Explain the logic of a **hierarchical model**
- Apply the **binomial distribution** to describe the sum of multiple Bernoulli trials
- Program a hierarchical binomial model in Stan
- Analyze secondary data using a hierarchical normal model (i.e., random-effect meta-analysis)

## Task List

1. Review the resources (slides and [notes][17])
2. Watch the lecture videos below (to be posted)
3. Complete the assigned readings
    * McElreath ch. 13.1, 13.2
    * [Gabry et al. (2019)][12]
    * [Gelman et al. (2020)][13]
4. Start working on Homework 3 (see instruction on Brightspace)

[12]: https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssa.12378
[13]: https://arxiv.org/abs/2011.01808
[17]: docs/05-hierarchical-models.html

## Lecture Videos

::: {.panel-tabset}

### Hierarchical Models

{{< video https://youtu.be/U_W2wsBRGNY >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "In terms of statistical inference, a binomial variable $Z$ with *N* = 10 and theta ($\\theta$) is equivalent to",
    answer = "10 binary variables that are exchangeable, and each has the same theta", 
    options = c("10 Bernoulli variables that are dependent, and each has the same theta", 
                "10 Bernoulli variables that are exchangeable, and each has a different theta",
                "10 binary variables that are dependent, and each has a different theta",
                "10 binary variables that are exchangeable, and each has the same theta"), 
    type = "radio"
)
```

### Partial Pooling

{{< video https://youtu.be/GLMXkzls0YA >}}

{{< video https://youtu.be/pRj3lyBu12o >}}

::: {.callout-note icon=false}

## Check your learning

Based on the notation in the slides, a Beta2(0.5, 6) distribution is the same as

```{r}
#| echo: false
#| results: asis
check_question(
    right = right_ans,
    wrong = "Consider how $\\mu$ and $\\kappa$ are defined in the Beta2 distribution",
    answer = "A Beta distribution with a = 3, b = 3", 
    options = c("A Beta distribution with a = 2, b = 6", 
                "A Beta distribution with a = 3, b = 3",
                "A Beta distribution with a = 4, b = 2"), 
    type = "radio"
)
```

:::

```{r}
#| echo: false
#| results: asis
practice_question(
    "A Gamma distribution is handy as a prior for kappa ($\\kappa$) because",
    answer = "A Gamma distribution is defined in the range 0 to infinity, which matches the range of kappa", 
    options = c("A Gamma distribution is defined in the range 0 to infinity, which matches the range of kappa", 
                "A Gamma distribution only has one parameter, which is easy to work with",
                "A Gamma distribution is a conjugate prior to the Beta model"), 
    type = "radio"
)
```

### Hierarchical Binomial

::: {.callout-tip icon="false" appearance="minimal"}
The following videos are from an older class, so the slides may look slightly different.
:::

::: {.callout-tip icon="false" appearance="minimal"}
For the first video, please skip to 04:12.
:::

{{< video https://youtu.be/GJcs8utKhBQ
    start="252"
>}}

{{< video https://youtu.be/GCpRvcrqQ9E >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "With shrinkage, the posterior distribution of individual-specific $\\theta_j$ is",
    answer = "pulled towards the center of the common distribution of theta", 
    options = c("pulled farther away from the center of the common distribution of theta", 
                "typically more concentrated due to borrowing information from other clusters",
                "A Gamma distribution is a conjugate prior to the Beta model"), 
    type = "radio"
)
```

### Hierarchical Normal

::: {.callout-tip icon="false" appearance="minimal"}
The following videos are from an older class, so the slides may look slightly different.
:::

{{< video https://youtu.be/GmB2rs-W1eE >}}

{{< video https://youtu.be/OC9oH5GRErU >}}

{{< video https://youtu.be/yQExiCLVcCk >}}

```{r}
#| echo: false
#| results: asis
practice_question(
    "In the hierarchical normal distribution discussed in the lecture, the treatment effect estimate in each study",
    answer = "has a different mean and a difference standard deviation", 
    options = c("has the same mean with the hierarchical model", 
                "has a different mean but the same standard deviation",
                "has a different mean and a difference standard deviation"), 
    type = "radio"
)
```

:::

## Slides

<div>

```{=html}
<iframe class="slide-deck" src="slides/hierarchical_models.html" width="525" height="350"></iframe>
```

</div>

[Link to HTML slides](/slides/hierarchical_models.qmd){target="_blank"}

:::

P.S.: If'd like to print the slides to PDF, follow <https://quarto.org/docs/presentations/revealjs/presenting.html#print-to-pdf>