---
title: "Exercise 1"
author: "Instructor of PSYC 573"
notes: modified from <https://quarto.org/docs/get-started/authoring/rstudio.html>
---

## Introduction

In this analysis, we will run a Bayesian analysis to estimate the death rate of patients diagnosed with AIDS in Australia. Let's start by loading the packages we'll use for the analysis.

```{r}
#| label: load-pkgs
#| code-summary: "Packages"
#| message: false

library(tidyverse)  # for data wrangling and visualization
library(knitr)      # for tables
library(posterior)  # for summarizing Bayesian analyses
```

<!-- Ex1: Install the `tidyverse` and the `posterior` packages so that they can be loaded above. Do not include the `install.packages()` statement in the qmd document. -->

```{r}
#| label: import-data
data(Aids2, package = "MASS")
```

We present the results of exploratory data analysis in @sec-eda and the regression model in @sec-model. See <!-- Ex2: Add citation for our textbook () here, using the "insert citation" function in RStudio. -->

## Exploratory data analysis {#sec-eda}

The data contains `{r} nrow(Aids2)` participants.

### Summary statistics

@tbl-state displays basic summary statistics.

```{r}
#| label: tbl-state
#| tbl-cap: "Patient status by state"

table(Aids2$state, Aids2$status) |>
  kable()
```

## Modeling {#sec-model}

We can fit a simple linear regression model of the form shown in @eq-slr.

<!-- \[Ex 3: Using LaTeX, add here the equation shown in this picture: https://wikimedia.org/api/rest_v1/media/math/render/svg/4211e3e7c3482573cdfbc0653d48a6279104c899\] -->

@tbl-lm shows the regression output for this model.

```{r}
#| label: tbl-lm
#| tbl-cap: "Summary of posterior distribution of death rate among patients diagnosed with AIDS in Australia"

prior_a <- 2
prior_b <- 2

posterior_a <- prior_a + sum(Aids2$status == "D")
posterior_b <- prior_b + sum(Aids2$status == "A")

posterior_draws <- rbeta(4000, shape1 = posterior_a, shape2 = posterior_b)

list(theta = posterior_draws) |>
  summarize_draws() |>
  kable(digits = 2)
```

@fig-histogram shows the prior and the posterior distributions of the death rate.

<!-- Ex4: Change the aspect ratio of the graph above to 1:1, and `fig-width` to 3.5 -->

```{r}
#| label: fig-histogram
#| fig-cap: "Histograms of prior and posterior distributions of death rate"
#| fig-subcap:
#|   - "Prior distribution"
#|   - "Posterior distribution"
#| layout-ncol: 2

prior_draws <- rbeta(4000, shape1 = prior_a, shape2 = prior_b)
ggplot(data.frame(theta = prior_draws), aes(x = theta)) +
  geom_histogram(binwidth = 0.03) +
  labs(title = "Prior", x = expression(theta)) +
  scale_x_continuous(limits = c(0, 1), oob = scales::oob_keep)

ggplot(data.frame(theta = posterior_draws), aes(x = theta)) +
  geom_histogram(binwidth = 0.01) +
  labs(title = "Posterior", x = expression(theta)) +
  scale_x_continuous(limits = c(0, 1), oob = scales::oob_keep)
```

<!-- Ex 5: Make the following section a callout box -->

## Results

The estimated death rate is `{r} round(mean(posterior_draws), 2)`, 95% CI [`{r} paste0(round(quantile(posterior_draws, c(.025, .975)), 2), collapse = ", ")`].

<!-- Ex 6: Change the YAML so that
a. It contains your name
b. Change the output to PDF (pdf or typst)
c. Add a table of content -->

## References {.unnumbered}
