---
title: "Exercise 10"
author: "Instructor of PSYC 573"
---

```{r}
#| message: false
library(here)
library(tidyverse)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 2)
library(rstan)
```

## Instruction

Answer questions 1-7 below. You can simply submit your responses in text, without using Quarto.

```{r}
#| code-fold: true
datfile <- here("usc-psyc573-notes/data", "marginalp.xlsx")
marginalp <- readxl::read_excel(datfile)
# Recode `Field` into a factor
marginalp <- marginalp |>
    # Filter out studies without any experiments
    filter(`Number of Experiments` >= 1) |>
    mutate(Field = factor(Field,
        labels = c(
            "Cognitive Psychology",
            "Developmental Psychology",
            "Social Psychology"
        )
    )) |>
    # Rename the outcome
    rename(marginal_p = `Marginals Yes/No`)
marginalp <- marginalp |>
  mutate(Year10 = (Year - 1970) / 10)
marginalp_cog <- filter(marginalp,
                        Field == "Cognitive Psychology")
```

The following fits the logistic model:

$$
\begin{aligned}
  \text{marginal\_p}_i & \sim \mathrm{Bern}(\mu_i)  \\
  \mathrm{logit}(\mu_i) & = \eta_i \\
  \eta_i & = \beta_0 + \beta_1 \text{Year10}_{i}
\end{aligned}
$$

where `marginal_p` is a binary variable indicating whether a paper contains any marginal $p$ values ($.05 < p \leq .10$), and `Year10` = (Year - 1970) / 10. For example, in the Year 2010, `Year10` = 4.

```{r}
#| code-fold: true
#| message: false
m1 <- brm(marginal_p ~ Year10,
    data = marginalp_cog,
    family = bernoulli(link = "logit"),
    prior = c(
        prior(student_t(4, 0, 1), class = "b"),
        prior(student_t(4, 0, 2.5), class = "Intercept")
    ),
    file = "ex10_m1"
)
m1
```

```{r}
#| code-fold: true
plot(
    conditional_effects(m1, prob = .90),
    points = TRUE,
    point_args = list(height = 0.01, width = 0.05, alpha = 0.05),
    plot = FALSE
)[[1]] +
    scale_x_continuous(
        breaks = 0:4,
        labels = c("1970", "1980", "1990", "2000", "2010")
    ) +
    xlab("Year")
```

Q1: Why is there no $\sigma$ parameter?

Q2: Based on the model estimated coefficients, what is the predicted log odds of Y (i.e., having a marginal $p$ value) for the year 2000?

Q3: What is the predicted probability of Y for the year 2000? [Note: Probability = exp(log odds) / (1 + exp(log odds))]

Q4: What is the predicted probability of Y for the year 2010?

Q5: The probability of Y in 2010 is ____ times the probability of Y in 2000.

Q6: The odds ratio can be computed as $\exp(\beta_1)$ = _______. Is this number the same as the probability ratio in Q5? Why or why not?

Q7: $\beta_1 / 4$ = _______. The difference in probability of Y from 2000 to 2010 is _______. 