---
title: "Exercise 3"
author: "Instructor of PSYC 573"
eval: false
---

```{r}
#| message: false
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
```

In this exercise, you will complete the analysis for a Bernoulli model in Stan. **Please make sure you change the YAML option to `eval: true`**.

## Data

You'll use the Therapeutic Touch data discussed in the lectures. Specifically, there are two participants: `S02`, who guessed the correct hand 2 times out of 10, and `S08`, who guessed the correct hand 8 times out of 10.

```{r}
# Data for S01 (1 out of 10)
data_s02 <- list(
    N = 10,
    y = c(1, 1, 0, 0, 0, 0, 0, 0, 0, 0)
)
# Data for S28 (8 out of 10)
data_s28 <- list(
    N = 10,
    y = c(1, 1, 1, 1, 1, 1, 1, 1, 0, 0)
)
```

Q1: Compile the attached "beta-bernoulli-pp.stan" file with the following code (Make sure the Stan file is in the same folder as this qmd):

```{stan}
#| output.var: bern_mod
#| file: "beta-bernoulli-pp.stan"
#| eval: false
```

```{r}
bern_mod <- cmdstan_model("beta-bernoulli-pp.stan")
```

Q2: As the model compiled successfully, the following performs MCMC sampling for participant `S02` to estimate the "ability" of guessing the correct hand (i.e., $\theta$). Repeat the sampling for participant `S28`.

```{r}
s02_fit <- bern_mod$sample(data_s02)
# Add code for `s28_fit`

```

Q3: Plot the posterior distribution of $\theta$ for the two participants.

```{r}
# First, extract theta
# Note: an MCMC draw is simply a simulated value of the parameter
#       from the posterior distribution
theta_s02 <- s02_fit$draws("theta")
mcmc_hist(theta_s02)
# For S28, try mcmc_dens()
theta_s28 <- s28_fit$draws("theta")

```

Q4: Use the following code to obtain the posterior distribution of the **difference** in $\theta$ between the two participants. Compute the probability that $\theta$ for participant `S28` is greater than that of participant `S02`.

```{r}
theta_diff <- theta_s28 - theta_s02
# Compute the probability that theta_s28 > theta_s02

```