---
title: "Exercise 4"
author: "Instructor of PSYC 573"
eval: false
---

```{r}
#| message: false
library(tidyverse)
library(cmdstanr)
# register_knitr_engine()
library(bayesplot)
```

In this exercise, we will perform some prior predictive checks for a hierarchical model to evaluate the sensibility of different priors. **Please make sure you change the YAML option to `eval: true`**.

## Data

You'll use the Therapeutic Touch example discussed in the lectures, but we don't need the data as we'll just focus on the prior part.

Q1: Copy the following Stan code and save it as a new file "hierarchical-binomial-prior.stan".

```{stan}
#| eval: false
#| output.var: hbin_mod
data {
  int<lower=0> J;  // number of clusters (e.g., studies, persons)
  array[J] int N;  // sample size for each cluster
  real prior_a;  // hyperparameter "a" for gamma
  real prior_b;  // hyperparameter "b" for gamma
}
parameters {
  // cluster-specific probabilities
  vector<lower=0, upper=1>[J] theta;
  real<lower=0, upper=1> mu;  // overall mean probability
  real<lower=0> kappa;        // overall concentration
}
model {
  // Priors
  theta ~ beta_proportion(mu, kappa);
  mu ~ beta(1.5, 1.5);      // weak prior
  kappa ~ gamma(prior_a, prior_b);
}
generated quantities {
  // Prior predictive
  array[J] int prior_ytilde = binomial_rng(N, theta);
}
```

Q2: Compile the Stan model using the code below (and modify the file path as needed).

```{r}
hbin_mod <- cmdstan_model("hierarchical-binomial-prior.stan")
```

Q3: The following shows the prior predictive distribution with `prior_a = .01` and `prior_b = .01`. In a few sentences, explain what the following two code blocks do.

<!-- Write down your response below -->

```{r}
prior_fit1 <- hbin_mod$sample(
    data = list(J = 28,
                N = rep(10, 28),
                prior_a = .01,
                prior_b = .01),
    seed = 1424,  # for reproducibility
    refresh = 1000
)
```

```{r}
ytilde1 <- prior_fit1$draws("prior_ytilde", format = "draws_matrix")
ppd_hist(ytilde1[sample.int(4000, size = 12), ],
         breaks = 0:10) +
    scale_x_continuous(breaks = 0:10) +
    labs(x = "Simulated counts of correct responses with N = 28")
```

Q4: Add code below to repeat the analyses in Q3, but use `prior_a = 2` and `prior_b = .1`.

```{r}
# insert R code
```

Q5: Comparing the plots in Q3 and Q4, how does the two different gamma priors affect the predicted data?

<!-- Insert your response for Q5 -->

Q6: Use the `bayesplot::ppd_stat()`{.r} function to compare the predicted standard deviation of the data. Please refer to the documentation of the function.

