---
title: "Exercise 6"
author: "Instructor of PSYC 573"
echo: false
format: pdf
---

The analyses in this exercise are from the example in the note "9 Multiple Predictors" (<https://marklhc.quarto.pub/psyc573-2024fall/docs/06b-multiple-predictors.html#conditional-effectssimple-slopes>)

```{r}
#| message: false
library(tidyverse)
library(brms)
options(brms.backend = "cmdstanr")
```

```{r}
#| include: false
waffle_divorce <- read_delim(  # read delimited files
    "https://raw.githubusercontent.com/rmcelreath/rethinking/master/data/WaffleDivorce.csv",
    delim = ";"
)
# Rescale Marriage and Divorce by dividing by 10
waffle_divorce$Marriage <- waffle_divorce$Marriage / 10
waffle_divorce$Divorce <- waffle_divorce$Divorce / 10
waffle_divorce$MedianAgeMarriage <- waffle_divorce$MedianAgeMarriage / 10
# Recode `South` to a factor variable
waffle_divorce$South <- factor(waffle_divorce$South,
    levels = c(0, 1),
    labels = c("non-south", "south")
)
# See data description at https://rdrr.io/github/rmcelreath/rethinking/man/WaffleDivorce.html
```

S = 0 for non-southern states; S = 1 for southern states

Consider the interaction model

$$
  \begin{aligned}
    D_i & \sim N(\mu_i, \sigma)  \\
    \mu_i & = \beta_0 + \beta_1 S_i + \beta_2 A_i + \beta_3 S_i \times A_i
  \end{aligned}
$$

## Q1

Express, in terms of the model parameters (e.g., $\beta_0$, $\beta_1$),

(a) the predicted divorce rate ($\mu$) for a southern state with `MedianAgeMarriage` = 2.5: _____________________

(b) the predicted $\mu$ for a non-southern state with `MedianAgeMarriage` = 2.5: _____________________

(c) the difference between (a) and (b): ____________________________

## Q2

```{r}
#| include: false
m_inter <- brm(
    Divorce ~ South * MedianAgeMarriage,
    data = waffle_divorce,
    prior = prior(normal(0, 2), class = "b") +
        prior(normal(0, 10), class = "b", coef = "Southsouth") +
        prior(normal(0, 10), class = "Intercept") +
        prior(student_t(4, 0, 3), class = "sigma"),
    seed = 941,
    file = "ex6_inter"
)
```

The following shows the estimated coefficients (from `brms`)

```{r}
fixef(m_inter) |>
    knitr::kable(digits = 2)
```

and the interaction plot:

```{r}
plot(
    conditional_effects(m_inter,
        effects = "MedianAgeMarriage",
        conditions = data.frame(South = c("south", "non-south"),
                                cond__ = c("South", "Non-South"))
    ),
    points = TRUE
)
```

Label $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, and $\sigma$ in the graph above (or describe where they are in your words).
