---
title: "Exercise 8"
author: "Instructor of PSYC 573"
eval: false
format: pdf
---

```{r}
library(dagitty)
```

## Instruction

There are four DAGs in the pictures attached to this exercise on Brightspace. For each DAG, identify (a) which path is a causal path from X to Y, (b) which variables need to be adjusted to satisfy the back-door criterion for the causal effect of X --> Y. Below is an example:

## Q1: Top Left (Example)

Causal paths are X --> Y. The paths (i) X <-- Z --> Y and (ii) X <-- Z <-- A --> Y are not causal. Conditioning on Z is sufficient to block all backdoor paths. 

```{r top-left}
dag1 <- dagitty("dag{ X -> Y; X <- Z -> Y; X <- Z <- A -> Y }")
# The part below is optional; it just matches the appearance in the questions
coordinates(dag1) <- list(x = c(X = 0, Z = 1, A = 2, Y = 2),
                          y = c(X = 0, Z = -1, A = -1, Y = 0))
plot(dag1)  # verify it's the same as the given graph
# Variables to be conditioned on
adjustmentSets(dag1, exposure = "X", outcome = "Y")
```

## Q2: Top Right

## Q3: Bottom Left

## Q4: Bottom Right