---
title: "HW 8 Template"
author: "Mark Lai"
date: "`r Sys.Date()`"
echo: false
format:
  gfm:
    toc: true
---

```{r}
#| message: false
library(here)
library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(cmdstanr)  # use two cores
library(posterior)
library(bayesplot)
```

# Research Question

> Is there difference in student evaluation rating for native and non-native English-speaking instructors?

# Variables

- `profevaluation`: evaluation rating of the instructor: 1 (very unsatisfactory) to 5 (excellent)
- `nonenglish`: 1 = non-native English speakers, 0 = native-English speakers

## Data Import

```{r}
beauty <- read_excel(here("data_files", "ProfEvaltnsBeautyPublic.xls"))
```

## Variable Summary

Table @tbl-summ-var shows the summary statistics of evaluationg ratings by groups.

```{r}
#| label: tbl-summ-var
#| tbl-cap: Descriptive statistics by groups
datasummary(profevaluation  * 
                (N + Mean + SD + Min + Max + Histogram) ~ 
                factor(nonenglish, labels = c("Native", "Non-Native")),
            data = beauty)
```

# Model

Let $Y$ = profevaluation, $G$ = nonenglish

Model:
$$
  \begin{aligned}
    Y_{i, G = 0} & \sim N(\mu_1, \sigma_1) \\
    Y_{i, G = 1} & \sim N(\mu_2, \sigma_2)
  \end{aligned}
$$

Prior:
$$
  \begin{aligned}
    \mu_1 & \sim N(3, 2) \\
    \mu_2 & \sim N(3, 2) \\
    \sigma_1 & \sim N^+(0, 2) \\
    \sigma_2 & \sim N^+(0, 2)
  \end{aligned}
$$

## Analysis

We used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups). 

```{r}
#| include: false
mod <- cmdstan_model(here("hw_templates/stan_code", "normal_2group.stan"))
# 1. form the data list for Stan
stan_dat <- with(
    beauty,
    list(
        N1 = sum(nonenglish == 0),
        N2 = sum(nonenglish == 1),
        y1 = profevaluation[which(nonenglish == 0)],
        y2 = profevaluation[which(nonenglish == 1)]
    )
)
# 2. Run Stan
fit <- mod$sample(
    stan_dat,
    seed = 1234 # for reproducibility
)
```

# Results

As shown in the rank histogram in @fig-rank-hist-fit below, the chains mixed well.

```{r}
#| label: fig-rank-hist-fit
#| fig-cap: Rank histogram of the posterior distributions of model parameters.
as_draws(fit) |>
    mcmc_rank_hist(pars = c("mu1", "mu2", "sigma1", "sigma2"))
```

@tbl-summ-fit shows the posterior distributions of $\mu_1$, $\mu_2$, $\sigma_1$, $\sigma_2$, and $\mu_2 - \mu_1$.

```{r}
#| label: tbl-summ-fit
#| tbl-cap: Posterior summary of the model parameters.
summ_fit <- fit$draws() |>
    subset_draws(variable = c("mu1", "mu2", "sigma1", "sigma2")) |>
    mutate_variables(`mu2 - mu1` = mu2 - mu1) |>
    summarise_draws()
knitr::kable(summ_fit, digits = 2)
```

The analysis showed that on average, non-native speaking instructors received a lower evaluation rating than native speaker instructors, with a posterior mean of `r round(summ_fit$mean[5], 2)` and a 90% CI of [`r round(summ_fit$q5[5], 2)`, `r round(summ_fit$q95[5], 2)`].
